### 2013-01-14 sections/conclusion.tex
updated nfql source code page
back to 6 pages

We presented
\ac{NFQL}\footnote{\texttt{\url{https://github.com/vbajpai/nfql}}}, an
efficient C implementation of the stream-based flow query language. The
language allows applying absolute or relative filters, aggregating flows
into groups, evaluating timing relationships among them, and merging them into
one collection. \ac{NFQL} can execute such complex queries in matter of
minutes, thereby expanding the scope of current flow record processing tools.
The conducted performance evaluations reveal that \ac{NFQL} is on par with
tools that support only absolute filters. SiLK, the only openly available
package that provides tools that are similar to the rest of \ac{NFQL}'s
processing pipeline appears faster. This is because it can optimize its
operations in favor of the limited set of comparisons that are only based on
equality, and its usage of a different file storage format. The evaluation
queries developed as a part of this research work may also develop into a more
general benchmark suite for flow query tools and platforms.

>>>>

We presented
\texttt{nfql}\footnote{\texttt{\url{http://nfql.vaibhavbajpai.com}}}, an
efficient C implementation of the stream-based flow query language, \ac{NFQL}.
The language allows applying absolute or relative filters, aggregating flows
into groups, evaluating timing relationships among them, and merging them into
one collection. \texttt{nfql} can execute such complex queries in matter of
minutes, thereby expanding the scope of current flow record processing tools.
The conducted performance evaluations reveal that \texttt{nfql} is on par with
tools that support only absolute filters. SiLK, the only openly available
package that provides tools that are similar to the rest of \texttt{nfql}'s
processing pipeline appears faster. This is because it can optimize its
operations in favor of the limited set of comparisons that are only based on
equality, and its usage of a different file storage format. The evaluation
queries developed as a part of this research work may also develop into a more
general benchmark suite for flow query tools and platforms.

### 2013-01-14 sections/evaluation.tex
back to 6 pages

It can be seen that the performance of the filter stage in \ac{NFQL} is
comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
lower ratios, but then SilK and \texttt{nfdump} also use their own file
format. \texttt{nfdump} appears to be significantly faster than the rest. This
is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. It comes
as a realization, that adding \texttt{lzo} compression will drastically
improve \ac{NFQL}'s filter performance. It is important to note, that all the
tools were single-threaded in this evaluation, and did not completely utilize
the $24$ available cores.

>>>>

It can be seen that the performance of the filter stage in \ac{NFQL} is
comparable to that of \texttt{flowtools} and SiLK. SiLK takes less time on
lower ratios, but then SilK and \texttt{nfdump} also use their own file
format. \texttt{nfdump} appears to be significantly faster than the rest. This
is because \texttt{nfdump} uses the \texttt{lzo} compression scheme. Adding
\texttt{lzo} compression will likely improve \ac{NFQL}'s filter
performance. Note that all the tools were single-threaded
in this evaluation, and did not completely utilize the $24$ available cores.

### 2013-01-14 sections/introduction.tex
back to 6 pages

NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information
export. NetFlow \cite{rfc3954} is a network protocol designed by Cisco
Systems, which allows routers to generate and export flow records to a
designated collector. \ac{IPFIX} \cite{rfc5101} on the other hand is an open
standard by IETF based on NetFlow v$9$. The novelty of the standard lies in
its ability to describe record formats at runtime using templates based on an
extensible and well-defined information model. The data transfer mechanism is
also simplistic and extensible by being unidirectional and transport protocol
agnostic.

The wide applicability of this approach is easily seen from the pervasive use
of flow records for a vibrant set of network analysis applications. For
instance, a survey by Sperotto \textit{et al.} \cite{sperotto:2010} gives an
overview of how network flow analysis can be used to detect Denial of Service
attacks, scan probing target systems and the target discovery phase of worms.
In addition, a survey by Callado \textit{et al.} \cite{callado:2009} also
lists behaviour analysis of Internet backbone traffic and general anomaly
detection.

Understanding intricate traffic patterns require sophisticated flow analysis
tools that can mine flow records for such a usage.  Unfortunately current
tools fail to deliver owing to their language design and simplistic filtering
methods.  We recently proposed a flow query language design
\cite{vmarinov:2009} that aims to cater to such needs.  It can process
flow records, aggregate them into groups, apply absolute or relative filters
and invoke Allen interval algebra rules \cite{fallen:1983}. The expressiveness
of the language can be seen from \cite{vperelman:2011} where the authors
formulate flow queries to identify flow signatures of popular applications.

Flowy \cite{kkanev:2010} was a first feature complete Python prototype of that
flow query language. Due to performance problems, it was superseeded by a
complete rewrite in C, called Flowy 2.0 \cite{jschauer:thesis:2011}. In this
paper, we introduce \ac{NFQL} which extends on Flowy 2.0, making it more
feature complete and optimizing its execution engine with crispier algorithms.
We show that this iteration of our work is able to scale to real-world sized
traces and has comparable execution times to contemporary flow analysis tools.

The paper is organized as follows. In section \ref{sec:relatedwork} we survey
the current state-of-the-art flow-processing tools and reason how \ac{NFQL} is
different from each one of them. In section \ref{sec:design} we describe the
flow query language by discussing each stage of the processing pipeline. In
section \ref{sec:implementation} we introduce the inner workings of \ac{NFQL}.
It begins by providing an overview of the \ac{NFQL} architecture and the
structure of the intermediate format used to exchange messages. The workflow of
the execution engine is described next and is supplemented by a number of
performance optimizations made to make the implementation scale. The section is
followed by performance evaluations comparing \ac{NFQL} against contemporary
flow processing tools alongwith with a discussion on its current limitation and
future outlook in section \ref{sec:evaluation}, with conclusion in section
\ref{sec:conclusion}.

\begin{figure*}[!t]
\centering
\includegraphics[width=0.9\linewidth]{nfql-pipeline}
\caption{NFQL Processing Pipeline \cite{vmarinov:2009}}
\label{fig:nfql-pipeline}
\end{figure*}

>>>>

NetFlow and \ac{IPFIX} are the two popular protocols for IP flow information
export. NetFlow \cite{rfc3954} is a network protocol designed by Cisco
Systems, which allows routers to generate and export flow records to a
designated collector. \ac{IPFIX} \cite{rfc5101} is an open standard defined by
the IETF based on NetFlow v$9$. The wide applicability of this approach is
easily seen from the pervasive use of flow records for a set of different
network analysis applications. For instance, a survey by Sperotto \textit{et
al.} \cite{sperotto:2010} gives an overview of how network flow analysis can
be used to detect intrusion attacks. In addition, a survey by Callado
\textit{et al.} \cite{callado:2009} lists behaviour analysis of Internet
backbone traffic and general anomaly detection.

Understanding intricate traffic patterns require sophisticated flow analysis
tools that can mine flow records for such a usage.  Unfortunately current
tools fail to deliver owing to their language design and simplistic filtering
methods.  We recently proposed a flow query language design
\cite{vmarinov:2009} that aims to cater to such needs.  It can process flow
records, aggregate them into groups, apply absolute or relative filters and
invoke Allen interval algebra rules \cite{fallen:1983} to merge group records.
The expressiveness of the language can be seen from \cite{vperelman:2011}
where the authors formulate flow queries to identify flow signatures of
popular applications.

Flowy \cite{kkanev:2010} was a first feature complete Python prototype of
\ac{NFQL}. Due to performance problems, it's execution engine was rewritten in
C, called Flowy 2.0 \cite{jschauer:thesis:2011}. In this paper, we introduce
\texttt{nfql}, which extends Flowy 2.0, making it more feature complete and
optimizing its execution engine with crispier algorithms.  We show that this
iteration of our work is able to scale to real-world sized traces and has
comparable execution times to contemporary flow analysis tools.

The paper is organized as follows. In Section \ref{sec:relatedwork} we survey
the current state-of-the-art of flow-processing tools and we reason how
\texttt{nfql} is different from each one of them. In Section \ref{sec:design}
we describe the flow query language by discussing each stage of the processing
pipeline. In Section \ref{sec:implementation} we provide an overview of the
\texttt{nfql} architecture and we describe the workflow of the execution
engine. A performance evaluations comparing \texttt{nfql} against contemporary
flow processing tools alongwith with a discussion on its current limitation
and future outlook is in section \ref{sec:evaluation}. We conclude the paper
in Section \ref{sec:conclusion}.

%\begin{figure}[ht]
  %\begin{minipage}[b]{0.5\linewidth}
    %\centering 
    %\includegraphics[scale=1]{nfql-pipeline}
    %\caption{Devebec’s Facade System}
    %\label{fig:devebec}
  %\end{minipage}	
  %\begin{minipage}[b]{0.5\linewidth}
    %\centering
    %\includegraphics[scale=1]{nfql-architecture}
    %\caption{MIT City Scanning Project}
    %\label{fig:mit}
  %\end{minipage}	
%\end{figure}

%\begin{figure}
        %\centering
        %\begin{subfigure}[b]{0.3\textwidth}
                %\centering
                %\includegraphics[width=\textwidth]{nfql-pipeline}
                %\caption{NFQL}
                %\label{fig:nfql-pipeline}
        %\end{subfigure}%
        %~ %add desired spacing between images, e. g. ~, \quad, \qquad etc.
          %%(or a blank line to force the subfigure onto a new line)
        %\begin{subfigure}[b]{0.3\textwidth}
                %\centering
                %\includegraphics[width=\textwidth]{nfql-architecture}
                %\caption{A tiger}
                %\label{fig:tiger}
        %\end{subfigure}
        %\caption{Pictures of animals}\label{fig:animals}
%\end{figure}

\begin{figure*}[!t]
  \subfloat{\includegraphics[width=0.7\textwidth]{nfql-pipeline}}
  \label{foo}
  \hfil
  \subfloat{\includegraphics[width=0.3\textwidth]{nfql-architecture}}
  \label{bar}
  \caption{NFQL Processing Pipeline \cite{vmarinov:2009} (left) and \texttt{nfql}
Architecture (right)}
  \label{fig:nfql-pipeline}
\end{figure*}

%\begin{figure*}[!t]
%\centering
%\includegraphics[width=0.9\linewidth]{nfql-pipeline}
%\caption{NFQL Processing Pipeline \cite{vmarinov:2009}}
%\label{fig:nfql-pipeline}
%\end{figure*}
